\subsection{Initial Results}
As described in Section~\ref{sec:proposed_approach}, we conducted a series of initial experiments to evaluate the performance of various machine learning models on the prediction of major oxide compositions from our \gls{libs} dataset.
These experiments aimed to provide a preliminary assessment of the models' performance.
All models were trained on the same preprocessed data using the Norm 3 preprocessing method described in Section~\ref{sec:norm3}.
This ensured that the models' performance could be evaluated under consistent and comparable conditions.

Table~\ref{tab:init_results} presents the results of these experiments, including the \gls{rmsep}, \gls{rmsecv}, standard deviation, standard deviation of the cross-validation, and the mean of these metrics for each model and oxide.
The primary metric of interest is the \gls{rmsecv}, as it provides a measure of the model's predictive accuracy during cross-validation, reflecting its generalization performance on unseen data.
Additionally, the standard deviation of the cross-validation offers insight into the variability of the model's performance across different folds, and the \gls{rmsep} indicates the model's predictive accuracy on a separate test set.

Ridge regression and \gls{svr} generally exhibit strong performance across different oxides, as evidenced by their relatively low \gls{rmsecv} values.
In contrast, \gls{enet}, \gls{ann}, and \gls{cnn} tend to perform the worst, with higher \gls{rmsecv} values.
Models such as \gls{rf} and \gls{gbr} exhibit moderate performance, often surpassing simpler linear models but not as consistently as \gls{svr}.

For $\ce{SiO2}$, \gls{svr} demonstrates the lowest \gls{rmsecv} of 4.908.
Ridge regression and \gls{rf} also perform well, with \gls{rmsecv} values of 5.005 and 5.304, respectively.
The standard deviations for these models are relatively low, indicating stable performance.

In the case of $\ce{TiO2}$, \gls{etr} exhibits the lowest \gls{rmsecv} of 0.427, signifying exceptional performance, with \gls{svr} and \gls{ngboost} also showing strong predictive capabilities, with \gls{rmsecv} values of 0.463 and 0.433, respectively.
he low standard deviation of the cross-validation for these models indicates consistent performance across folds.

For $\ce{Al2O3}$, \gls{ngboost} with an \gls{rmsecv} of 2.291 and ridge regression with an \gls{rmsecv} of 3.115 achieve relatively low \gls{rmsecv}, highlighting their effectiveness in predicting this oxide, with \gls{svr} also performing well with an \gls{rmsecv} of 2.700.
he standard deviation analysis reveals that \gls{ngboost} has the least variability, making it a more reliable model for this oxide.

\gls{svr} emerges as the best-performing model for $\ce{FeO_T}$ with the lowest \gls{rmsecv} of 2.847, while ridge regression with an \gls{rmsecv} of 3.385 and \gls{ngboost} with an \gls{rmsecv} of 3.561 also demonstrate competent performance.
The standard deviation values indicate that while \gls{svr} has lower variability, the other models also maintain reasonable consistency.

For $\ce{MgO}$, \gls{svr} exhibits the best performance with the lowest \gls{rmsecv} of 1.426, and ridge regression with an \gls{rmsecv} of 1.509 and \gls{ngboost} with an \gls{rmsecv} of 1.578 are also effective.
The standard deviation metrics suggest that these models are stable, with \gls{svr} showing the least variation.

\gls{gbr} shows the lowest \gls{rmsecv} of 1.468 for $\ce{CaO}$, indicating strong performance, with ridge regression having an \gls{rmsecv} of 1.513 and \gls{etr} having an \gls{rmsecv} of 1.503 also performing well.
Standard deviation values also highlight \gls{gbr} and \gls{etr} as having consistent performance.

\gls{etr} demonstrates the best performance for $\ce{Na2O}$ with the lowest \gls{rmsecv} of 1.028, while ridge regression with an \gls{rmsecv} of 1.221 and \gls{svr} with an \gls{rmsecv} of 1.096 also perform reasonably well.
The standard deviation of the cross-validation for \gls{etr} is particularly low, underscoring its robustness.

For $\ce{K2O}$, \gls{etr} again shows strong performance with the lowest \gls{rmsecv} of 0.681, with \gls{svr} having an \gls{rmsecv} of 0.690 and ridge regression having an \gls{rmsecv} of 0.668 also performing well.
The standard deviation analysis supports these findings, indicating reliable performance for these models.

To summarize, \gls{svr} stands out as the best overall model based on its average performance, with a mean \gls{rmsecv} of 1.958, closely followed by \gls{ngboost} with a mean \gls{rmsecv} of 2.017 and ridge regression with a mean \gls{rmsecv} of 2.111. \gls{svr}, ridge regression, and \gls{ngboost} are consistently among the top-performing models for multiple oxides.
Additionally, \gls{etr} performs exceptionally well for specific oxides such as $\ce{TiO2}$, $\ce{Na2O}$, and $\ce{K2O}$, suggesting that it may be a strong candidate for predicting these oxides.

\input{sections/results/init_results_table.tex}