\newpage

\input{sections/summary.tex}

\newpage

\begin{abstract}
This thesis advances the analysis of \gls{libs} data for predicting major oxide compositions in geological samples.
By integrating machine learning techniques and ensemble regression models, the study addresses challenges like high dimensionality, multicollinearity, and limited data availability.
Key innovations include the use of stacked generalization for improved model performance and an automated hyperparameter optimization framework.
The research contributes a comprehensive catalog of models and preprocessing techniques, and integrates findings into the \gls{pyhat} by the \gls{usgs}, enhancing its scientific capabilities.
This work aims to establish a robust foundation for future advancements in geochemical analysis and planetary exploration using \gls{libs} data.
\end{abstract}

\maketitle

\subsubsection*{Acknowledgements:} We would like to thank our supervisors Daniele Dell'Aglio and Juan Manuel Rodriguez for their guidance and support throughout this project.
We also thank our external supervisor Jens Frydenvang for his guidance as a domain expert from the ChemCam team, as well as volunteering his time to provide feedback on our work.
Furthermore, we extend our gratitude to Ryan B. Anderson and Travis S. Gabriel for their invaluable discussions regarding calibration and quantification based on \gls{libs} data. We also thank them for the opportunity to contribute to \gls{pyhat}.

\input{sections/introduction.tex}
\input{sections/related_work.tex}
\input{sections/problem_definition.tex}
\input{sections/background/index.tex}
\input{sections/baseline_replica.tex}
\input{sections/proposed_approach/proposed_approach.tex}
\input{sections/methodology.tex}
\input{sections/experiments/index.tex}
\input{sections/pyhat_contribution.tex}
\input{sections/conclusion.tex}
\input{sections/future_work.tex}
